---
title: "582-HW1"
author: "Abdulsamed Kağıt"
date: "18/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1 

Plot the following histogram diagrams for each season: 
a. Home Score(goals)
b. Away Score(goals)
c. Home Score(goals)– Away Score(goals)

To which probability distribution do home and away goals fitting well?  Does the distribution look like Poisson distribution? Calculate the expected number 
of games corresponding to each quantile (number of goals) with Poisson distribution by using sample means as distribution mean and plot these values on the histogram. Is this consistent with Poisson distribution claim? 

```{r Data Reading, message=FALSE, warning=FALSE, , echo=FALSE}
library(ggplot2)
library(dplyr)
library(yardstick)
library(broom)
library(data.table)
library(stringr)
library(tidyr)

match_2018 <- fread("C:/Users/Abdulsamed/Desktop/IE582/E0.csv")
match_2019 <- fread("C:/Users/Abdulsamed/Desktop/IE582/E1.csv")
match_2020 <- fread("C:/Users/Abdulsamed/Desktop/IE582/E2.csv")

match_2019 <- match_2019 %>% select(-Time)
match_2020 <- match_2020 %>% select(-Time)

```

### For Season 2018/2019

Required data is plotted in the histograms below and red line represents the fitted Poisson distribution.
Home Score: It can be seen that Poisson distribution fits well to the histogram of Home Score.  

```{r Home Score 2018/2019, message=FALSE, warning=FALSE}
expected_goal <- 0:6
lambda <- mean(match_2018$FTHG)
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2018)*dpois(expected_goal,lambda))

ggplot(match_2018) + geom_histogram(aes(x=FTHG),binwidth = 1,col="black") + labs(title="Season 2018/2019", x ="Home Goals", y = "Number of Games")+ theme_bw()+ 
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)

```
Away Score: Seems like Poisson distribution fits well to the histogram of Away Score too.  

```{r Away Score 2018/2019, message=FALSE, warning=FALSE}
lambda <- mean(match_2018$FTAG)
expected_goal <- 0:6
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2018)*dpois(expected_goal,lambda))

ggplot(match_2018,aes(x=FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2018/2019", x ="Away Goals", y = "Number of Games")+ theme_bw()+
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)
```
Home Score - Away Score: Histogram of score differences resembles Gaussian distribution rather than Poisson distribution.

```{r Home Score - Away Score 2018/2019, message=FALSE, warning=FALSE}
ggplot(match_2018,aes(x=FTHG-FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2018/2019", x ="Home Goals - Away Goals", y = "Number of Games")+ theme_bw() 
```

### For Season 2019/2020

Home Score: Poisson distribution fits well to the histogram of Home Score in 2019-2020 season too.

```{r Home Score 2019/2020, message=FALSE, warning=FALSE}
lambda <- mean(match_2019$FTHG)
expected_goal <- 0:8
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2019)*dpois(expected_goal,lambda))

ggplot(match_2019,aes(x=FTHG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2019/2020", x ="Home Goals", y = "Number of Games")+ theme_bw()+
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)
```
Away Score: Not surprisingly...

```{r Away Score 2019/2020, message=FALSE, warning=FALSE}
lambda <- mean(match_2019$FTAG)
expected_goal <- 0:9
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2019)*dpois(expected_goal,lambda))

ggplot(match_2019,aes(x=FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2019/2020", x ="Away Goals", y = "Number of Games")+ theme_bw()+ 
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)
```

Home Score - Away Score: This histogram still resembles Gaussian distribution but with lower variance comparing to previous season.

```{r Home Score - Away Score 2019/2020, message=FALSE, warning=FALSE}
ggplot(match_2019,aes(x=FTHG-FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2019/2020", x ="Home Goals - Away Goals", y = "Number of Games")+ theme_bw() 

```

### For Season 2020/2021

Home Score: Considering season 2020/2021 has not ended yet, this histogram is reasonably deviated from Poisson distribution.

```{r Home Score 2020/2021, message=FALSE, warning=FALSE}
lambda <- mean(match_2020$FTHG)
expected_goal <- 0:7
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2020)*dpois(expected_goal,lambda))

ggplot(match_2020,aes(x=FTHG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2020/2021", x ="Home Goals", y = "Number of Games")+ theme_bw()+
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)
```

Away Score: Note that if all 3 seasons were combined and analyzed together, the histogram of Home Score and Away Score would fit to the Poisson distribution better. Sample size is not large enough for this season yet. 

```{r Away Score 2020/2021, message=FALSE, warning=FALSE}
lambda <- mean(match_2020$FTHG)
expected_goal <- 0:7
expected_goal <- data.frame(x=expected_goal, y=nrow(match_2020)*dpois(expected_goal,lambda))

ggplot(match_2020,aes(x=FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2020/2021", x ="Away Goals", y = "Number of Games")+ theme_bw()+
geom_line(data = expected_goal, aes(x=x,y=y), color = "red", lwd = 1)

```

Home Score - Away Score:

```{r Home Score - Away Score 2020/2021, message=FALSE, warning=FALSE}
ggplot(match_2020,aes(x=FTHG-FTAG)) + geom_histogram(binwidth = 1,col="black") + labs(title="Season 2020/2021", x ="Home Goals - Away Goals", y = "Number of Games")+ theme_bw() 

```

## Task 2

Select at least 4 bookmakers for this task.

1)Calculate the P(home win), P(tie) and P(away win) by P(x) = 1/odd.

2)Then calculate these probabilities again using normalization formula at “Odds and Probabilities” part for each bookmarker.

3) First construct a plot of P(home win) – P(away win) on x-axis and P (tie) on y-axis with first probability calculation; then plot the actual probabilities calculated using the results.

4)Do this for each bookmaker separately. Comment on if there is a bias in odds representing the probabilities? Name the x and y axes accordingly. Write the name of bookmaker at the top of each plot.

#### Necessary Data Manipulations 

Data of 3 seasons are filtered and merged. Necessary calculations are performed as such:

```{r Task2, message=FALSE, warning=FALSE}

match_2018_filtered <- match_2018 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR)
match_2019_filtered <- match_2019 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR)
match_2020_filtered <- match_2020 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR)
match_combined <- bind_rows(match_2018_filtered,match_2019_filtered, match_2020_filtered)

match_combined<-cbind( match_combined[,1] ,apply(match_combined[,2:13],c(1,2), function(x) { 1/x}), match_combined[,c(14,15)] )
match_combined <- match_combined %>% mutate(B365_sum = apply(match_combined[,2:4],1, FUN = sum ))
match_combined <- match_combined %>% mutate(BW_sum = apply(match_combined[,5:7],1, FUN = sum ))
match_combined <- match_combined %>% mutate(IW_sum = apply(match_combined[,8:10],1, FUN = sum ))
match_combined <- match_combined %>% mutate(PS_sum = apply(match_combined[,11:13],1, FUN = sum ))

match_combined <- match_combined %>% mutate(B365H_Normal =B365H/B365_sum , B365D_Normal =B365D/B365_sum , B365A_Normal= B365A/B365_sum )
match_combined <- match_combined %>% mutate(BWH_Normal =BWH/BW_sum , BWD_Normal =BWD /BW_sum , BWA_Normal= BWA/BW_sum )
match_combined <- match_combined %>% mutate(IWH_Normal =IWH/IW_sum , IWD_Normal =IWD /IW_sum , IWA_Normal= IWA/IW_sum )
match_combined <- match_combined %>% mutate(PSH_Normal =PSH/PS_sum , PSD_Normal =PSD /PS_sum , PSA_Normal= PSA/PS_sum )
match_combined <-match_combined %>% select(- BW_sum, -IW_sum, - PS_sum,-B365_sum)

```

#### Plot for B365

Estimated probabilities are calculated by separating P(home win) – P(away win) values into 21 bins. Lower and Upper limit for these bins are determined as (-0.821, 0.899) respectively. Number of games ended as "Draw" within these bins are divided by total number of games corresponding to the bins, hence estimated probability of Draws are estimated and plotted to the histogram. Obviously estimated probabilities in intervals (-0.047,0.039] and (0.469,0.555] are higher than proposed by bookmaker B365. 

```{r B365 , message=FALSE, warning=FALSE}
match_B365 <- match_combined %>% select(FTR, B365H_Normal, B365D_Normal,B365A_Normal, HR,AR )
match_B365 <- match_B365 %>% mutate(H_A_interval = cut( (match_B365$B365H_Normal-match_B365$B365A_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_B365, aes(x= (B365H_Normal-B365A_Normal), y= B365D_Normal, col=H_A_interval))+
labs(title="B365 Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_B365_draw_count <- match_B365 %>% group_by(H_A_interval) %>% count(FTR)
match_B365_draw_count <- match_B365_draw_count %>%rename(draw_count = n)
match_B365_draw_count <- match_B365_draw_count %>% filter(FTR =="D")
match_B365_draw_count<- match_B365_draw_count %>% select(-FTR)
  
match_B365_count <-match_B365 %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_B365_count <- full_join(match_B365_count,match_B365_draw_count, by="H_A_interval")  
match_B365_count <- match_B365_count %>% replace_na(list(draw_count=0))

match_B365_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_B365, aes(x= (B365H_Normal-B365A_Normal), y= B365D_Normal, col=H_A_interval))+
  labs(title="B365 Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+ 
  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_B365_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)

```

#### Plot for BW

Estimated probabilities in intervals (0.297,0.383] and (0.641,0.727] are higher than proposed by bookmaker BW. 

```{r BW , message=FALSE, warning=FALSE}
match_BW <- match_combined %>% select(FTR, BWH_Normal, BWD_Normal,BWA_Normal, HR,AR )
match_BW <- match_BW %>% mutate(H_A_interval = cut( (match_BW$BWH_Normal-match_BW$BWA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_BW, aes(x= (BWH_Normal-BWA_Normal), y= BWD_Normal, col=H_A_interval))+
  labs(title="BW Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_BW_draw_count <- match_BW %>% group_by(H_A_interval) %>% count(FTR)
match_BW_draw_count <- match_BW_draw_count %>%rename(draw_count = n)
match_BW_draw_count <- match_BW_draw_count %>% filter(FTR =="D")
match_BW_draw_count<- match_BW_draw_count %>% select(-FTR)

match_BW_count <-match_BW %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_BW_count <- full_join(match_BW_count,match_BW_draw_count, by="H_A_interval")  
match_BW_count <- match_BW_count %>% replace_na(list(draw_count=0))

match_BW_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_BW, aes(x= (BWH_Normal-BWA_Normal), y= BWD_Normal, col=H_A_interval))+
  labs(title="BW Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_BW_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)
```


#### Plot for IW

Estimated probabilities in intervals (-0.047,0.039,] , (0.125,0.211] and (0.297,0.383] are higher than proposed by bookmaker IW. 

```{r IW , message=FALSE, warning=FALSE}
match_IW <- match_combined %>% select(FTR, IWH_Normal, IWD_Normal,IWA_Normal, HR,AR )
match_IW <- match_IW %>% mutate(H_A_interval = cut( (match_IW$IWH_Normal-match_IW$IWA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_IW, aes(x= (IWH_Normal-IWA_Normal), y= IWD_Normal, col=H_A_interval))+
  labs(title="IW Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_IW_draw_count <- match_IW %>% group_by(H_A_interval) %>% count(FTR)
match_IW_draw_count <- match_IW_draw_count %>%rename(draw_count = n)
match_IW_draw_count <- match_IW_draw_count %>% filter(FTR =="D")
match_IW_draw_count<- match_IW_draw_count %>% select(-FTR)

match_IW_count <-match_BW %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_IW_count <- full_join(match_IW_count,match_IW_draw_count, by="H_A_interval")  
match_IW_count <- match_IW_count %>% replace_na(list(draw_count=0))

match_IW_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_IW, aes(x= (IWH_Normal-IWA_Normal), y= IWD_Normal, col=H_A_interval))+
  labs(title="IW Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_IW_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)


```

#### Plot for PS

Estimated probabilities in intervals (-0.047,0.039,] and (0.125,0.211] are higher than proposed by bookmaker PS. 

```{r PS , message=FALSE, warning=FALSE}
match_PS <- match_combined %>% select(FTR, PSH_Normal, PSD_Normal,PSA_Normal, HR,AR )
match_PS <- match_PS %>% mutate(H_A_interval = cut( (match_PS$PSH_Normal-match_PS$PSA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_PS, aes(x= (PSH_Normal-PSA_Normal), y= PSD_Normal, col=H_A_interval))+
  labs(title="PS Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_PS_draw_count <- match_PS %>% group_by(H_A_interval) %>% count(FTR)
match_PS_draw_count <- match_PS_draw_count %>%rename(draw_count = n)
match_PS_draw_count <- match_PS_draw_count %>% filter(FTR =="D")
match_PS_draw_count<- match_PS_draw_count %>% select(-FTR)

match_PS_count <-match_PS %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_PS_count <- full_join(match_PS_count,match_PS_draw_count, by="H_A_interval")  
match_PS_count <- match_PS_count %>% replace_na(list(draw_count=0))

match_PS_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_PS, aes(x= (PSH_Normal-PSA_Normal), y= PSD_Normal, col=H_A_interval))+
  labs(title="PS Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_PS_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)
```

## Task 3

There can be some events during the matches that create noise in the outcomes. To be more specific, let’s consider a specific case. Bookings can affect the game result. A red card in the first few minutes of a game can change the outcome of the match drastically. Playing with few players is always a disadvantage for the teams.

Matches with red cards are discarded and all operations in task 2 are performed all over again to see if there is any significant change in the observations made.

```{r Task3, message=FALSE, warning=FALSE}
match_2018_filtered <- match_2018 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR) %>% filter( HR ==0, AR == 0)
match_2019_filtered <- match_2019 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR)%>% filter( HR ==0, AR == 0)
match_2020_filtered <- match_2020 %>% select(FTR, B365H,B365D,B365A,BWH,BWD,BWA,IWH,IWD,IWA,PSH,PSD,PSA,HR,AR)%>% filter( HR ==0, AR == 0)
match_combined <- bind_rows(match_2018_filtered,match_2019_filtered, match_2020_filtered)

match_combined<-cbind( match_combined[,1] ,apply(match_combined[,2:13],c(1,2), function(x) { 1/x}), match_combined[,c(14,15)] )
match_combined <- match_combined %>% mutate(B365_sum = apply(match_combined[,2:4],1, FUN = sum ))
match_combined <- match_combined %>% mutate(BW_sum = apply(match_combined[,5:7],1, FUN = sum ))
match_combined <- match_combined %>% mutate(IW_sum = apply(match_combined[,8:10],1, FUN = sum ))
match_combined <- match_combined %>% mutate(PS_sum = apply(match_combined[,11:13],1, FUN = sum ))

match_combined <- match_combined %>% mutate(B365H_Normal =B365H/B365_sum , B365D_Normal =B365D/B365_sum , B365A_Normal= B365A/B365_sum )
match_combined <- match_combined %>% mutate(BWH_Normal =BWH/BW_sum , BWD_Normal =BWD /BW_sum , BWA_Normal= BWA/BW_sum )
match_combined <- match_combined %>% mutate(IWH_Normal =IWH/IW_sum , IWD_Normal =IWD /IW_sum , IWA_Normal= IWA/IW_sum )
match_combined <- match_combined %>% mutate(PSH_Normal =PSH/PS_sum , PSD_Normal =PSD /PS_sum , PSA_Normal= PSA/PS_sum )
match_combined <-match_combined %>% select(- BW_sum, -IW_sum, - PS_sum,-B365_sum)
```

#### Plot for B365

Estimated probabilities in intervals (-0.047,0.039] and (0.469,0.555] are higher than proposed by bookmaker B365. No change occurred. 

```{r B365-2 , message=FALSE, warning=FALSE}
match_B365 <- match_combined %>% select(FTR, B365H_Normal, B365D_Normal,B365A_Normal, HR,AR )
match_B365 <- match_B365 %>% mutate(H_A_interval = cut( (match_B365$B365H_Normal-match_B365$B365A_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_B365, aes(x= (B365H_Normal-B365A_Normal), y= B365D_Normal, col=H_A_interval))+
  labs(title="B365 Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_B365_draw_count <- match_B365 %>% group_by(H_A_interval) %>% count(FTR)
match_B365_draw_count <- match_B365_draw_count %>%rename(draw_count = n)
match_B365_draw_count <- match_B365_draw_count %>% filter(FTR =="D")
match_B365_draw_count<- match_B365_draw_count %>% select(-FTR)

match_B365_count <-match_B365 %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_B365_count <- full_join(match_B365_count,match_B365_draw_count, by="H_A_interval")  
match_B365_count <- match_B365_count %>% replace_na(list(draw_count=0))

match_B365_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_B365, aes(x= (B365H_Normal-B365A_Normal), y= B365D_Normal, col=H_A_interval))+
  labs(title="B365 Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_B365_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)

```

#### Plot for BW

Estimated probabilities in intervals (0.297,0.383] and (0.641,0.727] are higher than proposed by bookmaker BW. No change occurred.

```{r BW-2 , message=FALSE, warning=FALSE}
match_BW <- match_combined %>% select(FTR, BWH_Normal, BWD_Normal,BWA_Normal, HR,AR )
match_BW <- match_BW %>% mutate(H_A_interval = cut( (match_BW$BWH_Normal-match_BW$BWA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_BW, aes(x= (BWH_Normal-BWA_Normal), y= BWD_Normal, col=H_A_interval))+
  labs(title="BW Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_BW_draw_count <- match_BW %>% group_by(H_A_interval) %>% count(FTR)
match_BW_draw_count <- match_BW_draw_count %>%rename(draw_count = n)
match_BW_draw_count <- match_BW_draw_count %>% filter(FTR =="D")
match_BW_draw_count<- match_BW_draw_count %>% select(-FTR)

match_BW_count <-match_BW %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_BW_count <- full_join(match_BW_count,match_BW_draw_count, by="H_A_interval")  
match_BW_count <- match_BW_count %>% replace_na(list(draw_count=0))

match_BW_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_BW, aes(x= (BWH_Normal-BWA_Normal), y= BWD_Normal, col=H_A_interval))+
  labs(title="BW Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_BW_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)

```

#### Plot for IW

Estimated probabilities in intervals (-0.047,0.039,] , (0.125,0.211] and (0.297,0.383] are higher than proposed by bookmaker IW. No change occurred. 

```{r IW-2 , message=FALSE, warning=FALSE}

match_IW <- match_combined %>% select(FTR, IWH_Normal, IWD_Normal,IWA_Normal, HR,AR )
match_IW <- match_IW %>% mutate(H_A_interval = cut( (match_IW$IWH_Normal-match_IW$IWA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_IW, aes(x= (IWH_Normal-IWA_Normal), y= IWD_Normal, col=H_A_interval))+
  labs(title="IW Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_IW_draw_count <- match_IW %>% group_by(H_A_interval) %>% count(FTR)
match_IW_draw_count <- match_IW_draw_count %>%rename(draw_count = n)
match_IW_draw_count <- match_IW_draw_count %>% filter(FTR =="D")
match_IW_draw_count<- match_IW_draw_count %>% select(-FTR)

match_IW_count <-match_BW %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_IW_count <- full_join(match_IW_count,match_IW_draw_count, by="H_A_interval")  
match_IW_count <- match_IW_count %>% replace_na(list(draw_count=0))

match_IW_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_IW, aes(x= (IWH_Normal-IWA_Normal), y= IWD_Normal, col=H_A_interval))+
  labs(title="IW Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_IW_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)

```

#### Plot for PS

Estimated probabilities in intervals (-0.047,0.039] and (0.125,0.211] are higher than proposed by bookmaker PS. No change occurred.

```{r PS-2 , message=FALSE, warning=FALSE}
match_PS <- match_combined %>% select(FTR, PSH_Normal, PSD_Normal,PSA_Normal, HR,AR )
match_PS <- match_PS %>% mutate(H_A_interval = cut( (match_PS$PSH_Normal-match_PS$PSA_Normal),breaks = seq(-0.821,0.899,length.out=21)))


ggplot(match_PS, aes(x= (PSH_Normal-PSA_Normal), y= PSD_Normal, col=H_A_interval))+
  labs(title="PS Probability Calculations", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") 

match_PS_draw_count <- match_PS %>% group_by(H_A_interval) %>% count(FTR)
match_PS_draw_count <- match_PS_draw_count %>%rename(draw_count = n)
match_PS_draw_count <- match_PS_draw_count %>% filter(FTR =="D")
match_PS_draw_count<- match_PS_draw_count %>% select(-FTR)

match_PS_count <-match_PS %>% group_by(H_A_interval) %>% summarise( H_A_interval = H_A_interval[1], total_count = n() ) 

match_PS_count <- full_join(match_PS_count,match_PS_draw_count, by="H_A_interval")  
match_PS_count <- match_PS_count %>% replace_na(list(draw_count=0))

match_PS_count$H_A_interval<-seq(-0.7757368,0.8537368,length.out=20) #assign mid points for plotting

ggplot(match_PS, aes(x= (PSH_Normal-PSA_Normal), y= PSD_Normal, col=H_A_interval))+
  labs(title="PS Probability Calculations with Estimated Probability of Draws", x =" P(home win) – P(away win)", y = " P (tie)")+  scale_colour_discrete("P(Home)-P(Away)")+ theme_bw() +
  geom_point() + geom_smooth(color="black",se=FALSE,size=2,linetype = "dashed") + 
  geom_col(data=match_PS_count, aes(x=H_A_interval, y= (draw_count/total_count)),col="red",alpha=0.3)

```









